\chapter*{Acknowledgements}
I have been more successful and satisfied with my Ph.D. career than I had dared 
to hope when I started, and I can't imagine it happening anywhere but Penn.  
This has not so much to do with The City of Brotherly Love, the ``greene 
Country Towne'', the so-called American Paris, despite its prideworthy grit and 
soul, its food trucks, diners, and BYOBs, its parks and squares filled with 
dogs and children, its brick and rainbow row houses, its Chinatown and Italian 
Market, Schuykill River Trail and Wissahickon.
Philadelphia has made a lasting impression on me, but it is the revolving cast 
of characters around me these last five years that have made all the 
difference:

I'd like to thank Kostas Daniilidis for convincing me to come to Penn with his 
generous hospitality while introducing me to the GRASP lab.  I have continued 
to enjoy his hospitality and skillful diplomacy throughout my career, including  
while
heading my thesis committee.  I'd also like to thank Jianbo Shi for further 
persuading me to come to Penn, and more importantly for lighting the computer 
vision fire inside of me; something I will keep for the rest of my life.
I am also indebted to the rest of my thesis committee: C.J. Taylor for his 
tenacious and thorough following of both my Written Preliminary Examination II 
and this thesis, which produced thoughtful comments and corrections.  Thanks 
also to David Forsyth for his warmth, enthusiasm for my research, and inspiring 
high-level vision.

It is hard to imagine working with any other advisor than Ben Taskar.  I am 
grateful for his unwavering optimism and confidence that things will work out.  
His hands-on approach, hard work ethic, and adherence to the highest possible 
quality and rigor are now an indelible part of my own approach to research.  I 
also cherish our friendship: cracking jokes, playing soccer, enjoying 
conferences, and knowing that he has my best interests at heart in work and 
life.

Instrumental in my formative Ph.D. years was Timothee Cour, with whom I cut my 
teeth on my first serious vision projects, who showed me the way of MATLAB 
hacking, and how to cut quick to the heart of a problem.  Timothee, Alex Toshev 
and Katerina Fragkiadaki are all equal parts colleagues, confidants and 
comrades, at equal ease proposing and critiquing research ideas, enjoying 
drinks, and exploring new cities and countries to the fullest.  Thanks to David 
Weiss for his enthusiasm, deep discussions about life and work, and easy access 
to cats; best of luck in the future.

I can't thank Mike Felker and Charity Payne enough---Mike has always made me 
feel that someone is personally looking out for me in the Ph.D. program;  
Charity makes all administrative tasks seem effortless.


Thanks to BUGS for many enjoyable reading groups and outings---I have 
especially fond memories of the Poconos retreat and NIPS workshops (special 
credit due to the masterful organizer Alex Kulesza). Thanks to the rest of the 
GRASP lab for giving me a home-away-from-home.  Thanks to the Automatons, the 
FC Inter Penn All Stars, and Penn-pickup-soccer for giving me something to look 
forward to every week.  Finally, thanks to Yuri for keeping me grounded to the 
important things in life, especially laughter.


